Bioverse: Origins of Life

### 2023-08-05 ms.tex
Update on Overleaf.

@@

>>>>

\section{Bayesian Analysis}
It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true. 
This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the occurrence rate of the planet type required by the hypothesis $\eta_i$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding

\begin{align}
\label{eqn:pbio}
\pdf(bio|H_i) = \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
\end{align}

However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
To obtain this we can use Bayes' theorem, which yields 
\begin{align}
\pdf(H_i|bio) = \pdf(bio|H_i) \pdf(H_i) / \pdf(bio).
\end{align}
Putting everything together, we arrive at
\begin{align}
\label{eqn:posterior}
\pdf(H_i|bio) = \frac{\pdf(H_i)}{\pdf(bio)} \times \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
\end{align}

Here, $\pdf(bio)$ is the prior probability of detecting the biosignature.
Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
\begin{align}
\pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \eta_i \times \pdf_\mathrm{det, i}.
\end{align}
$\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$.
If we assume that all OoL hypotheses are a priori equally probable, we can treat $\frac{\pdf(H_i)}{\pdf(bio)}$ in Eq.~\ref{eqn:posterior} as a normalization constant.

In the following, we discuss the impact of the remaining variables $\pdf_\mathrm{abio}, \eta_i, $ and $\pdf_\mathrm{det, i}$ on $\pdf(H_i|bio)$.

\subsection{Probability of Abiogenesis $\pdf_\mathrm{abio}$}
Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
\begin{align}
\pdf_\mathrm{abio} = 1 - \exp(-\lambda t).
\end{align}
~\\
...

\subsection{Planet occurrence rate $\eta_i$}
\subsection{Detection probability $\pdf_\mathrm{det, i}$}

@@

''Which natural processes best explain how living matter spontaneously appears from nonliving matter?''~\citep{Malaterre2022}
%WE consider life ``as we know it''
